
## Regression Models: Course Project

# Does the type of transmission affects the fuel consumption?

#### Sagi Greenstine

## Executive Summary

The Motor Trend magazine staff is interested in exploring the relationship between a **set of variables** at a data set of a collection of cars and **miles per gallon (MPG) (outcome)**.  
They are especially interested in the answers to the following questions:

1. Is an automatic or manual transmission better for MPG?  
2. How significant MPG difference between automatic and manual transmissions?

This report is based on the ```mtcars``` dataset and after the following data analysis, I can do these conclusions:


Brief. 
Roughly the equivalent of 2 pages or less for the main text. 
Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. 
The appendix can only include figures.
Include a first paragraph executive summary.

Did the student interpret the coefficients correctly?
Did the student do some exploratory data analyses?
Did the student fit multiple models and detail their strategy for model selection?
Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
Did the student do a residual plot and some diagnostics?
Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?

## Data Analysis

### Exploratory Data Analysis

The ```mtcars``` dataset is a data frame with 32 observations on 11 variables.  
1. First of all, I built the pairs plot for the ```mtcars``` dataset variables and found that there are strong correlations between the variables: **hp, cyl, disp, wt, mpg** (see **Plot 1** in **Appendix**).  
2. I built the boxplot of MPG values by transmissions type and found that the MPG of a **manual transmission** significantly higher than the MPG of an **automatic transmission** (see **Plot 2** in **Appendix**).


### Simple Linear Regression

### Multiple Regression

* Simple Linear Regression - lm(mpg ~ am, data = mtcars)
* Multiple Regression - lm(mpg ~ wt + qsec + am, data=mtcars)

### Results

## Conclusions

## Appendix

### Exploratory Data Analysis

#### The ```mtcars``` dataset
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).
Source: Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.  

There is a some view on the data:

```{r viewdata}
head(mtcars)
```

The data frame with 32 observations on 11 variables.  
[, 1]	mpg	Miles/(US) gallon  
[, 2]	cyl	Number of cylinders  
[, 3]	disp	Displacement (cu.in.)  
[, 4]	hp	Gross horsepower  
[, 5]	drat	Rear axle ratio  
[, 6]	wt	Weight (lb/1000)  
[, 7]	qsec	1/4 mile time  
[, 8]	vs	V/S  
[, 9]	am	Transmission (0 = automatic, 1 = manual)  
[,10]	gear	Number of forward gears  
[,11]	carb	Number of carburetors  

Now I'll make some exploratory analysis of the data.

**Plot 1: The Scatterplot Matrix for ```mtcars``` Dataset Variables**

I'll build the scatterplot by using the ```gclus``` package:

```{r applot1, echo=FALSE, message=FALSE}
library(gclus)
data <- mtcars[c(1:11)] 
datarows <- abs(cor(data))
datacols <- dmat.color(datarows)
dorder <- order.single(datarows) 
cpairs(data, dorder, panel.colors=datacols, gap=.5, main="mtcars Variables Correlations" )
```

We can see that the strongest correlations exist between the variables **hp, cyl, disp, wt, mpg**.

**Plot 2: The Boxplot MPG by Transmission Type** 

```{r applot2, echo=FALSE, message=FALSE}
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
boxplot(mpg ~ am, data = mtcars, col = (c("green","red")), ylab = "MPG", xlab = "Transmission Type")
```

Here we can see that the MPG of a **manual transmission** significantly higher than the MPG of an **automatic transmission**.
