---
output: pdf_document
---

## Regression Models: Course Project

# Does the type of transmission affects the fuel consumption?

#### Sagi Greenstine

## Executive Summary

The Motor Trend magazine staff is interested in exploring the relationship between a **set of variables** at a data set of a collection of cars and a **miles per gallon (MPG) (outcome)**.  
They are especially interested in the answers to the following questions:  
1. Is an automatic or manual transmission better for MPG?  
2. How significant MPG difference between automatic and manual transmissions?  
This report is based on the ```mtcars``` dataset and after the following data analysis, I can do these conclusions:  
1. The manual transmission is better for MPG than the automatic.  
2. There is a signficiant difference between the MPG means of manual and automatic transmissions. In terms of MPG means, the difference between the two types of transmission is 7.24 mpgs (in favor of the manual transmission). On average, manual transmission has 1.55 mpgs more than automatic transmission.

## Data Analysis

### Exploratory Data Analysis

The ```mtcars``` dataset is a data frame with 32 observations on 11 variables.  
1. First of all, I built the pairs plot for the ```mtcars``` dataset variables and found that there are strong correlations between the variables: **hp, cyl, disp, wt, mpg** (see **Plot 1** in the **Appendix**).  
2. I built the boxplot of MPG values by transmissions type and found that the MPG of a **manual transmission** significantly higher than the MPG of an **automatic transmission** (see **Plot 2** in the **Appendix**).

Now, let's make the regression analysis.

### Simple Linear Regression
There is a simple linear regression for variables mpg and am.
```{r linregr}
model1 <-lm(mpg~am, data = mtcars)
summary(model1)
```

The adjusted R-squared is 0.3385, it means that our model only explains 33.85% of the regression variance.  
There are several other predictors that should be taken into account to see whether any one of them plays a bigger role in the model.  
Let's build the multiple regression model.

### Multiple Regression
We see that the strongest correlations exist between the variables **hp, cyl, disp, wt, mpg**. These variables will be used in our multiple regression model.
```{r mregr}
model2 <- lm(mpg~am + hp + cyl + disp + wt, data = mtcars)
summary(model2)
anova(model1, model2)
```

Based on the p-value, I can reject the null hypothesis and claim that the multiple model is significantly different from the simple model.  
Let's check the residuals for any signs of non-normality and examine the residuals versus fitted values plot for any signs of heteroskedasticity.

```{r res, fig.width=6, fig.height=6}
par(mfrow = c(2,2))
plot(model2)
```
  
**Residuals vs. Fitted**: the points are randomly scattered on the plot that indicates on the independence condition.  
**Normal Q-Q**: the points mostly fall on the line, and it indicates the normal distribution of the residuals.  
**Scale-Location**: the points scattered in a constant band pattern, and it that indicates the constant variance.  
**Residuals vs Leverage**: there are points in the top of the plot that may indicate values of increased leverage of outliers.

## Conclusions
- The residuals are homoscedastic and normally distributed.  
- By the multiple regression model I can see that the multiple R-squared value is 0.855, i.e. 85.5% of the regression variance can be explained by the chosen variables. I can conclude that these variables are confounding in the relationship between the **am** and **mpg*.  
- The manual transmission is better for MPG than the automatic.  
- There is a signficiant difference between the MPG means of manual and automatic transmissions: on average, manual transmission has 1.55 mpgs more than automatic transmission.  


## Appendix

### Exploratory Data Analysis

#### The ```mtcars``` dataset
The data frame with 32 observations on 11 variables (see the [description](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)).  
There is a some view on the data:

```{r viewdata}
head(mtcars)
```

Now I'll make some exploratory analysis of the data.

**Plot 1: The Scatterplot Matrix for ```mtcars``` Dataset Variables**

I'll build the scatterplot by using the ```gclus``` package:

```{r applot1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
library(gclus)
data <- mtcars[c(1:11)] 
datarows <- abs(cor(data))
datacols <- dmat.color(datarows)
dorder <- order.single(datarows) 
cpairs(data, dorder, panel.colors=datacols, gap=.5, main="mtcars Variables Correlations" )
```

We can see that the strongest correlations exist between the variables **hp, cyl, disp, wt, mpg**.  
Now let's see the MPG difference between **automatic** and **manual** transmissions:

**Plot 2: The Boxplot MPG by Transmission Type** 

```{r applot2, echo=FALSE, message=FALSE, fig.width=5, fig.height=5}
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
boxplot(mpg ~ am, data = mtcars, col = (c("red","green")), ylab = "MPG", xlab = "Transmission Type")
```

```{r mpgmeans}
aggregate(mpg~am, data = mtcars, mean)
```

Here we can see that the MPG mean of a **manual transmission** higher than the MPG mean of an **automatic transmission** by 7.24 mpgs.  
To find out whether it's a significant difference, I have to perform a following t-test:
```{r ttest}
AutoTrans <- mtcars[mtcars$am == "Automatic",]
ManualTrans <- mtcars[mtcars$am == "Manual",]
t.test(AutoTrans$mpg, ManualTrans$mpg)
```

The p-value is 0.00137, I can reject the null hypothesis and claim that there is a signficiant difference between the MPG means of manual and automatic transmissions.
